Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-07 paper/paper.tex
changes

\noindent
\tname{} only reports a rule if the \nids\ (\eg{}, Suricata)
instantiated with that rule is able to capture the traffic passed on
input. Consequently, by construction, a rule that uses this set of
options captures the malicious traffic. Unfortunately, this rule is
potentially overspecified. Informally, the rule includes ``too many''
characteristics of the malicious traffic used to define it and could
potentially result in false negatives. In this particular case, note
that the option \CodeIn{windows: 64} is not included in the golden
rule.

Second, \tname{} uses negative (\ie{}, benign) traffic to address the
overfitting issue. \tname{} searches for alternative rules that
preserves the invariant that a rule should capture the positive
traffic but not capture the negative traffic. It uses the rule
obtained in the previous step to bootstrap the search and generates
various rules containing subsets of these three options.

Finally, \tname{} ranks the inferred rules according to heuristics
extracted from a database of existing \tname{} rules. One example
heuristic is \Fix{...informa um exemplo concreto que
  vc. usaram...}. Instead of traffic, this step looks for the
structure of existing rules. The list below shows the rule options for
the top-5 rules \tname{} produces for this attack.

@@

@@

Active reconnaissance is a method used by malicious individuals to
collection information about a computing system to determine potential
vulnerabilities. Active reconnaissance is a preliminary step towards
an actual attack. The most popular technique is to ``scan'' open ports
in a host. A Ping Scan is a very common type of port scan based and it
is based on the ICMP protocol. \Mar{@Lucas, acima falamos de
  porta. Abaixo nao falamos de porta, apenas de maquina. Vc. pode ser
  mais especifico? ->}
  \Luc{De fato estamos falando de duas coisas diferentes, mas ambas são exemplos de 'active reconnaissance'. Primeiramente citamos o port scanning, q tem o objetivo de encontrar portas abertas em uma maquina especifica. Depois falamos do ping scan, q tem o objetivo de encontrar ip's ativos numa rede.} To scan a network, an attacker sends a several
ICMP Echo requests addressed to a range of IP addresses in the network
and checks which ones respond to the requests. This allow the attacker
to know which machines are active in the network, then it can choose
one of them to attack.

@@

Figure 4 shows a Suricata rule that captures a ping scan implemented
with NMAP~\cite{netmap}, a network discovery and security auditing
open-source tool. The relevant options are in bold---the option
\CodeIn{dsize: 0} indicates that the packet has no bytes in the
payload and the option \CodeIn{itype: 8} indicates that the packet to
be captured are ICMP Echo requests.

Again, to synthesize rules for a Ping Scan, \tname{} analyzes the
positive traffic and extracts the following rule options:

@@

\Fix{...show rules...}

@@

\tname{} is a technique to synthesize signature-based rules for
\nids~(\eg{}, Snort~\cite{snort} and Suricata~\cite{suricata}). The
goal of \tname\ is to create rules that captures \emph{only} the
malicious messages.

@@

\vspace{1ex}
\noindent\textbf{Overview.}~To create rules, \tname\ leverages (1)
benign and malicious network traffic and (2) rules from public
rulesets. Figure~\ref{fig:overview} shows the workflow of \tname{} as
a pipeline of three components, which appear numbered in the figure.
The inputs of the technique appear at the bottom of the figure whereas
the output appear at the right. \tname\ takes on input malicious and
benign traffic and a set of rules from public rulesets and produces on
output a list of plausible rules to isolate the malicious traffic.

>>>>

\noindent
\tname{} only reports a rule if the \nids\ instantiated with that rule
is able to capture the traffic passed on input. Consequently, by
construction, a rule that uses this set of options captures the
malicious traffic. Unfortunately, this rule is potentially
overspecified. Informally, the rule includes ``too specific''
characteristics of the malicious traffic and could potentially result
in false negatives. For example, in this particular case, note that
the option \CodeIn{windows: 64} is not included in the golden rule
from Figure~\ref{fig:synflood-example}.

Second, \tname{} uses negative (\ie{}, benign) traffic to mitigate the
overfitting issue. In this step, \tname{} searches for alternative
rules that preserves the invarianxt that a rule should capture the
positive traffic but not capture the negative traffic. It uses the
rule obtained in the previous step, containing three options, to
bootstrap the search for rules. The search generates rules containing
subsets of the options from the seed rule.

%% \Luc{Ataque "JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit"
%% . O payload do pacote é o seguinte:
%% \begin{figure}[H]
%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-rule.suricata}
%%   \label{fig:rule-example}
%% \end{figure}

%% A partir disso produzimos uma regra com os seguintes contents:

%% \begin{figure}[H]
%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-contents-splitted.suricata}
%%   \label{fig:rule-example}
%% \end{figure}

%% Depois da etapa da minimização, as top 5 regras são:

%% \begin{figure}[H]
%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-top5-rules.suricata}
%%   \label{fig:rule-example}
%% \end{figure}

%% A regra de ouro é:

%% \begin{figure}[H]
%%   \lstinputlisting[language=C,numbers=none,keywords={dsize,itype}]{adaptor-golden-rule.suricata}
%%   \label{fig:rule-example}
%% \end{figure}

Finally, \tname{} ranks the inferred rules according to heuristics
extracted from a database of existing rules. One example heuristic
function we used to rank the generated rules was \Fix{...informa um
  exemplo concreto que vc. usaram...}. The list below shows the rule
options for the top-5 rules \tname{} produces for this attack.

@@

To sum up, \tname{} uses the messages that manifest an attack in the
first step (to create an overspecific seed rule), benign traffic in
the second step (to create a set of plausible rules from the seed
rule), and existing rules from popular sources in the third step (to
rank the generated rules).

@@

%% \Mar{@Lucas, acima falamos de porta. Abaixo nao falamos de
%%   porta, apenas de maquina. Vc. pode ser mais especifico? ->} \Luc{De
%%   fato estamos falando de duas coisas diferentes, mas ambas são
%%   exemplos de 'active reconnaissance'. Primeiramente citamos o port
%%   scanning, q tem o objetivo de encontrar portas abertas em uma
%%   maquina especifica. Depois falamos do ping scan, q tem o objetivo de
%%   encontrar ip's ativos numa rede.}

Active reconnaissance is a method used by malicious individuals to
collect information about a computing system to determine potential
vulnerabilities. Active reconnaissance is a preliminary step towards
an actual attack. For example, Ping Scan is a very common type of
attack that scans the network for IPs of accessible hosts. An attacker
sends several ICMP Echo requests to a range of IP addresses in the
network and checks which ones respond to the requests. This allow the
attacker to know which machines are active in the network. Then (s)he
can choose one of them to attack.

@@

Figure 4 shows a Suricata rule that captures a reproduction of a Ping
Scan attack available from NMAP~\cite{netmap}, a network discovery and
security auditing open-source tool. The relevant options are in
bold---the option \CodeIn{dsize: 0} indicates that the packet has no
bytes in the payload and the option \CodeIn{itype: 8} indicates that
the packet to be captured are ICMP Echo requests.

Again, to synthesize rules for a Ping Scan, \tname{} analyzes the
positive malicious traffic and extracts the following rule options:

@@

\Fix{...listem 5 primeiras regras neste caso...}

@@

\tname{} is a technique to synthesize rules for signature-based
\nids~(\eg{}, Snort~\cite{snort} and Suricata~\cite{suricata}). The
goal of \tname\ is to create rules that capture \emph{only} the
malicious traffic.

@@

\vspace{1ex}
\noindent\textbf{Overview.}~To create rules, \tname\ leverages (1)
benign and malicious network traffic and (2) rules from public
rulesets. Figure~\ref{fig:overview} shows the workflow of \tname{} as
a pipeline of three components, which appear numbered in the figure.
The inputs of the technique appear at the bottom of the figure whereas
the output appears at the right. \tname\ takes on input malicious and
benign traffic and a set of rules from public rulesets and produces on
output a list of plausible rules to isolate the malicious traffic.
